{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-09-01 thesis/related.tex
relwork relaxed mem checkers

Other checkers target specific programming languages' concurrency models and/or thread communication APIs.
Relacy \cite{relacy} is a header-only C++ SCT library for verifying synchronization primitives.
It instruments the existing C++ atomics API,
although requires custom annotations for other types of memory accesses,
and notably includes relaxed memory nondeterminism in its concurrency model.
Inspect \cite{inspect} uses a static alias analysis to instrument and preempt all memory accesses to potentially-shared data.
SPIN \cite{spin} tests algorithms defined in the PROMELA domain-specific language,
instruments every memory access, and
specializes in verifying synchronization primitives such as RCU \cite{rcu}.
%
D\'{e}j\`{a} Fu \cite{dejafu} is a SCT
for the Haskell language,
whose strong type system guarantees that thread communication be confined to trusted, type-safe APIs.
%to a trusted API that implements internal synchronization to preserve type safety.
%Supporting both abstraction reduction and STM,
It instruments these interfaces (STM among them)
to check for deadlocks or nondeterministic behaviour in general,
which either may arise despite the static no-data-race guarantee.
The Rust language \cite{rust-language}
provides a more C++-like type system that also
%, which I contributed to personally,
achieves the same guarantee, although I know of no existing SCT for it yet.

@@

Several other recent advances extend DPOR to new concurrency models,
beyond the shared-memory-threading model outlined in \sect{\ref{sec:landslide-dpor}}.
TransDPOR \cite{transdpor} provides extra domain-specific reduction for message-passing actor programs
by exploiting the fact that the dependency relation is transitive in the absence of shared state.
The $R^4$ algorithm \cite{r4} (corresponding to the R4 checker mentioned above)
extends DPOR to event-driven programs by separating the notion of enabled events from that of multiple threads.
DPOR for TSO and PSO \cite{tsopso}
extends the concurrency model to include store buffer reordering,
which can expose bugs not even possible in the strong consistency model
such as discussed in \sect{\ref{sec:tm-warpzone-relaxed}}.
It also introduced a heuristic {\em buffer bounding} technique, analogous to ICB,
to mitigate the corresponding increase in state space size.
MCR was recently extended to support relaxed memory models likewise \cite{mcr-tsopso}.
Along the same lines, this work's Chapter~\ref{chap:tm}
extended DPOR's concurrency model for HTM,
and proposed two new reduction strategies, one sound and one heuristic,
to find new types of equivalences.
% TODO
% effective stateless model checking for c/c++ concurrency - popl 18

>>>>

Other checkers target specific programming languages' concurrency models and/or thread communication APIs.
Inspect \cite{inspect} uses a static alias analysis to instrument and preempt all memory accesses to potentially-shared data.
SPIN \cite{spin} tests algorithms defined in the PROMELA domain-specific language,
instruments every memory access, and
specializes in verifying synchronization primitives such as RCU \cite{rcu}.
%
D\'{e}j\`{a} Fu \cite{dejafu} is a SCT
for the Haskell language,
whose strong type system guarantees that thread communication be confined to trusted, type-safe APIs.
%to a trusted API that implements internal synchronization to preserve type safety.
%Supporting both abstraction reduction and STM,
It instruments these interfaces (STM among them)
to check for deadlocks or nondeterministic behaviour in general,
which either may arise despite the static no-data-race guarantee.
The Rust language \cite{rust-language}
provides a more C++-like type system that also
%, which I contributed to personally,
achieves the same guarantee, although I know of no existing SCT for it yet.

The problem of relaxed memory nondeterminism alone has inspired the creation of several new SCTs in the past few years.
Relacy \cite{relacy}, a header-only C++ SCT library for checking synchronization primitives,
was the first to broach this field,
although requires custom annotations for non-atomic memory accesses
and
%(according to later citing papers) % cdschecker
does not fully model all possible relaxed memory behaviours.
%designed for verifying synchronization primitives,
CDSChecker \cite{cdschecker} extends DPOR with a {\em reads-from} relation
to capture most of the C++11 memory model's new behaviours.
Nidhugg \cite{nidhugg} is a SCT for TSO and PSO which instruments LLVM abstract assembly,
% i don't actually know what the difference is #overlyhonestmethods
although does not yet support the C++11 memory model.
%% actually it doesn't -- they say they take the Source DPOR worse version. im confus?
% and extends the Optimal DPOR algorithm (discussed below) to include store buffer nondeterminism.
rInspect \cite{tsopso}
%models TSO and PSO differently, using shadow threads, and
offers further heuristic state space reduction using buffer bounding (described below).
RCMC \cite{rcmc} models a ``repaired'' version of the C++11 memory model known as RC11 \cite{rc11},
and professes to achieve the best state space reduction to date.
These tools each use various heuristics to account for spin-wait loops,
ranging from delay bounding \cite{bpor} to a rigid rewrite rule,
and provide only limited support so far for read-modify-write atomics
(at best, supporting them by introducing % oops
some redundant exploration).
%
No relaxed-memory SCT has yet proposed a satisfactory model for the ``thin-air'' problem \cite{sully-thesis},
which can cause state space cycles in a way not yet well-understood and remains future work.
%and notably includes relaxed memory nondeterminism in its concurrency model.
They also identify all data races
% TODO check if ther'es not a better section reference
(by the C++ definition rather than \sect{\ref{sec:quicksand-soundness}}'s)
as bugs immediately, rather than checking them for benign or buggy outcomes.
All of these tools are notably open-source -- an encouraging recent trend in the field.

@@

Several other recent advances extend DPOR to new concurrency models,
beyond the shared-memory-threading model outlined in \sect{\ref{sec:landslide-dpor}}.
TransDPOR \cite{transdpor} provides extra domain-specific reduction for message-passing actor programs
by exploiting the fact that the dependency relation is transitive in the absence of shared state.
The $R^4$ algorithm \cite{r4} (corresponding to the R4 checker mentioned above)
extends DPOR to event-driven programs by separating the notion of enabled events from that of multiple threads.
DPOR for TSO and PSO \cite{tsopso}
extends the concurrency model
using {\em shadow threads}, which interleave with traditional threads to represent store buffer nondeterminism,
which can expose bugs not even possible in the strong consistency model
such as discussed in \sect{\ref{sec:tm-warpzone-relaxed}}.
It also introduced a heuristic {\em buffer bounding} technique, analogous to ICB,
to mitigate the corresponding increase in state space size.
The same year, Nidhugg \cite{nidhugg} proposed a DPOR extension to account for TSO and PSO
using {\em chronological traces}.
MCR was recently extended to support relaxed memory models likewise \cite{mcr-tsopso}.
Just this year, RCMC \cite{rcmc} proposed to replace the interleaving model entirely with {\em execution graphs},
offering further reduction still.
Somewhat analogously for HTM, this work's Chapter~\ref{chap:tm}
extended DPOR's concurrency model to include failure injection,
% not "new" -- abstraction reduction is the same as always
and proposed three reduction strategies, one sound and two heuristic,
%to make some progress up the exponential mountain.
to keep state spaces manageable.
